A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-07-22 paper/paper.tex
paper: cooking some plans for restructuring

Type theory is a flexible formalism supporting different models. Some models are based on topological spaces, where equality is interpreted as path, and substitutivity of equality as transport~\citep{kapulkin21}. These models support the univalence principle stating that equality of types mimics equivalence of types, leading to the development of Homotopy Type Theory~\citep{hottbook}. In type theory, types are organized in a hierarchy of universes written $\Type_m$ for $m$ a natural number. Main types in type theory are the types of dependent pairs, written $\Sigma a : A.\,B(a)$, the types of dependent functions, written $\Pi a:A.\,B(a)$, for $A$ a type and $B(a)$ a type dependent on the inhabitant $a$ of $A$, and the type of propositional equality, written $t = u$. We assume our type theory to also include a distinguished singleton type, written $\unittype$, and with inhabitant $\unitpoint$, the type of boolean values, called $\textsf{bool}$, and the type of natural numbers. We also write $\hd$ and $\tl$ the projections of dependent pairs, and $\refl$ for reflexivity. Logical propositions being types themselves, we use $\Pi$ to represent universal quantification and $\Sigma$ to represent existential quantification.

@@

@@

Like in \ref{tab:coh}, we leave implicit in \ref{fulltab:coh}, the use of the isomorphism between the type of equalities of dependent pairs in a $\Sigma$-type with the $\Sigma$-type of equalities of the components of the pair, as well as the isomorphism between equality of functions with point-wise equality. Contrary to the ETT case, functional extensionality does not hold, so we assume it. However, the requirement of functional extensionality disappears if $\nu$ is finite. As for basic groupoid properties of equalities, which no longer hold definitionally, we do not mention them for simplicity.

>>>>

Type theory is a flexible formalism supporting different models. Some models are based on topological spaces, where equality is interpreted as path, and substitutivity of equality as transport~\citep{kapulkin21}. These models support the univalence principle stating that equality of types mimics equivalence of types, leading to the development of Homotopy Type Theory~\citep{hottbook}. In type theory, types are organized in a hierarchy of universes written $\Type_m$ for $m$ a natural number. Main types in type theory are the types of dependent pairs, written $\Sigma a : A.\,B(a)$, the types of dependent functions, written $\Pi a:A.\,B(a)$, for $A$ a type and $B(a)$ a type dependent on the inhabitant $a$ of $A$, and the type of propositional equality, written $t = u$. We assume our type theory to also include a distinguished singleton type, written $\unittype$, and with inhabitant $\unitpoint$, the type of boolean values, called $\textsf{bool}$, and the type of natural numbers. We also assume that our type theory includes the coinductive type of infinite tuples. We also write $\hd$ and $\tl$ the projections of dependent pairs, and $\refl$ for reflexivity. Logical propositions being types themselves, we use $\Pi$ to represent universal quantification and $\Sigma$ to represent existential quantification.

@@

% 5. The "mathematical" construction (and its intuition)
%   leaving equality and inequality side conditions implicit
%   in particular, this means:
%   - assuming equality reflection
%   - equality as in observational type theory
%   - assuming termination of recursive constructions
% 6. Formal details
% - making substitution of equals by equals explicit
% - observational properties of equality
% - the theory of inequality proofs
% - the well-foundedness
% - the formalization tricks
%   - implicit arguments
%   - refine-style vs tactic-style

@@

Contrary to the ETT case, functional extensionality does not hold, so we assume it. However, the requirement of functional extensionality disappears if $\nu$ is finite. As for basic groupoid properties of equalities, which no longer hold definitionally, we do not mention them for simplicity.
